---
title: "Quadratic Programming"
author: "Nischal Subedi"
date: "4/22/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


In this module, I will use existing optimization packages to solve quadratic programming (QP) problems. Recall the classical form of QP:
$$
\min _{x} \frac{1}{2} x^{T} P x+q^{T} x
$$
subject to $\quad G x \preceq h$
$$
A x=b
$$
Consider the following example:
$$
\begin{array}{l}
\min _{x, y \in \mathbb{R}} \frac{1}{2} x^{2}+3 x+4 y \\
\text { subject to } \quad x+4 y \geq 18 \\
2 x+5 y \leq 90 \\
3 x+4 y \leq 80 \\
x, y \geq 0
\end{array}
$$
We can then write this example equivalently as the classical form using vectors/matrices:
$$
\min _{x, y \in \mathbb{R}} \frac{1}{2}\left(\begin{array}{l}
x \\
y
\end{array}\right)^{T}\left(\begin{array}{ll}
1 & 0 \\
0 & 0
\end{array}\right)\left(\begin{array}{l}
x \\
y
\end{array}\right)+\left(\begin{array}{l}
3 \\
4
\end{array}\right)^{T}\left(\begin{array}{l}
x \\
y
\end{array}\right)
$$
$$
\text { subject to }\left(\begin{array}{cc}
-1 & -4 \\
2 & 5 \\
3 & 4 \\
-1 & 0 \\
0 & -1
\end{array}\right)\left(\begin{array}{c}
x \\
y
\end{array}\right) \leq\left(\begin{array}{c}
-18 \\
90 \\
80 \\
0 \\
0
\end{array}\right)
$$
We next use the R quadprog package  to solve the QP example above. Note that RMarkdown also allows you to run Python code by first installing **reticulate** package.




```{r}
library(quadprog)
library(Matrix)


# now we define the matrices/vectors based on the QP defined in the R documentation

Dmat=matrix(c(1,0,0,0),nrow=2,ncol=2)

Dvec=-c(3,4)
b=c(18,-90,-80,0,0)
Amat=t(matrix(c(1,-2,-3,1,0,4,-5,-4,0,1),nrow=5,ncol=2))

Dpd=nearPD(Dmat)$mat#this ensures that the D needs to be positive definite, if D is not positive definite then solve.OP will give error.

sol<-solve.QP(Dpd,Dvec,Amat,b,meq=0) # The first meq constraint are treated as equality constraint. Since, we do not have such constrant, we treat them as 0.

# now lets print out the optimal solution

sol$solution
sol$value


```







